I am grateful to so many people for all their support through my Ph.D. life. First,
I would like to say thanks to my parents who have always been there for me. I could
only see them once a year but thanks to technology I got to talk to them almost every day
and they could not be more encouraging. My brother has supported me throughout
my efforts as well, he is probably the main reason that I thought that graduate
school would be a great place for me. He always discussed my research with me and
kept me focused.

I have to thank my advisors Professor John Lafferty and Professor Hank Hoffmann for helping me so much in every aspect. 
John has been incredible advisor to me. He has taught me so much about machine 
learning and how to spot good problems to work on. We had been looking at different 
problems for a while until we met Hank who had some very interesting problems in 
computer architecture. I remember I was trying out different models until John just randomly 
pointed out some Bayesian models which have good properties and that model worked like magic.
I always noticed John never fits real world problems into machine learning 
but always looks at it the other way round. He has always been very kind and helpful to 
me and he's a rockstar of an advisor. Everything I know about computer systems comes from 
my interactions with Hank.
He's one of the most patient, hard-working and kind person I have known. He is
relentless in his efforts. One of my papers with him got us really bad reviews and 
I felt like I should just give up on systems research because its really hard to 
satisfy all the comments that reviewers told us. But, he reassured me that the work was good and 
the paper was finally selected. He has attended so many of my bad trial presentations
and always suggested changes which made big difference later in my final talks.
I am always a nervous speaker and he has given me a lot of time to practice with him 
and great advises from his previous experiences on how to cope with the pressure.
In terms of our work, Hank is a genius a extracting out machine learning problems
from seemingly regular and non-obvious systems problems. And clearly, it has helped us a lot.
He never made any of his students feel like we were working under him but rather with him.
He has contributed a lot in my thesis with not just discussion time but also running actual 
algorithms and writing a lot in the papers which is not very common in advisors. Again, John and Hank,
I thank you so much from the bottom of my heart. I could not have asked for a better people 
to work with. 

I must thank Professor Shan Lu for agreeing to be on my thesis committee and helping me improve my work. 
Shan has a knack for good writing and good presentation. Shan has pre-reviewed many of our 
papers and her comments have improved our paper drastically. I have enjoyed her questions 
during my candidacy exam and thesis defense. They are really good questions which are educational for the audience.

I should thank Professor Lazlo Babai for always being helpful. He taught us discrete math in first year. His class was
great, perhaps the best class I have ever taken. He's very devoted towards his student's success. Whenever we 
saw each other in the department he would make sure I had a good progress and offered help. I will always remember 
him as a great teacher. 

I must thank Bob, Collin and Phil for being super helpful techstaff. I have asked them for help for the most 
embarrassing questions and they have never refused or delayed. Some of my problems have been very specific but they would 
always find time to help me out. I still can't forget when my hard drive crashed on my computer, 
Bob spent over a week to make sure that I can get certain things out of it. I can easily say that my life at 
UChicago was made 50\% easier because of excellent techstaff in the department. They are indispensable.

Finally, it goes without saying how helpful my friends Qinqing, Liwen and Jiajun have been all this time.
I have shared my worst and best moments with them. Qinqing has taught me great coding practices and she has 
always been around for me to discuss my research with her. We have read so many papers together. She made me feel home 
when I missed my home the most. Liwen, Jiajun and I worked a lot during courses and
they have only been the best. They have helped my sail through many courses and have been all ears for any research related question.
They are very smart people.
