\section{Conclusion}
\label{sec:conclusion}
This paper has presented \SYSTEMLEO{}, a system capable of learning
Pareto-optimal power and performance tradeoffs for an application
running on a configurable system.  \SYSTEMLEO{} combines some of the best
features of both online and offline learning approaches.  Offline,
\SYSTEMLEO{} acquires knowledge about a range of application behaviors.
Online, \SYSTEMLEO{} quickly matches the observed behavior of a new
application to previously seen behavior from other applications to
produce highly accurate estimates of performance and power.  We have
implemented \SYSTEMLEO{}, made the source code available, and tested it
on a real system with 25 different applications exhibiting a range of
behaviors.  Across all applications, \SYSTEMLEO{} achieves greater than
97\% accuracy in its performance and power estimations despite only
sampling less than 2\% of the possible configuration space for an
application it has never seen before.  These estimations are then used
to allocate resources and save energy.  \SYSTEMLEO{} produces energy
savings within 6\% of optimal while purely Offline or Online
approaches are both over 24\% of optimal.  \SYSTEMLEO{}'s learning
framework represents a promising approach to help generalize resource
allocation in energy limited computing environments and could be used
in conjunction with other control techniques to help develop a
\emph{self-aware} computing system
\cite{Hoffmann2012,1508273,1333571,1516538,Kephardt2005,laddaga1999}.
